### 2019-02-05 thesis_report/results.tex
fixed zigzag figures

% -*- root: ./report.tex -*-
\chapter{Results \& Discussion}\label{chap:results}
		In this chapter, we show how some properties of the system change by the introduction of a zigzag geometry.
		Additionally, we attempt to isolate the mechanism behind the improved properties by estimating the gap size through the quasiclassical model introduced in chapter \ref{chap:methods}.

@@

			\begin{figure}[!htb]
			\centering
			\includegraphics[width=0.55\columnwidth]{figures/bandstructures}
			\caption{Figure of the bandstuctures corresponding to the system in Fig.~\ref{fig:zigzag_methods}.
			Blue lines correspond to $\phi=0$, $B=0$, red lines to $\phi=\pi$, $B = 1$.
			Thre three subplots are for different amplitudes of zigzag: (a) $z_y=0$, (b) $z_y=\frac{W}{4}$, and (c) $z_y=\frac{W}{2}$, where $W=200$\si{\nm} is the junction width.
			We observe that once there are no more straight trajectories inside the junction (when $z_y=\frac{W}{2}$) the spectrum becomes most insensitive to the momentum $k_x$.
			The value of the remaining parameters are $\alpha=20$ \si{\milli \eV \nm}, $g=26$, $\meff=0.02$\si{\electronmass}, $\mu=20$ \si{\milli \eV}, $\Delta=1$ \si{\milli \eV}.
			\label{fig:bandstuctures}}
			\end{figure}

			\comment{We calculate the bandstructure for varying amount of zigzag.}
			We implement a tight-binding version of the model described in section \ref{sec:physical_picture}.
			In figure~\ref{fig:bandstuctures}, the bandstructure for saw-toothed zigzag systems [Fig.~\ref{fig:zigzag_methods}] with varying amplitude is displayed.
			Because the zigzag has to be modelled using a supercell, we see a folded bandstructure.
			For comparison's sake, we take the same cell size for the ribbon structure, so the bandstructure folds in the same way.

			\comment{The band structures show that modulation of the geometry increases the gap by an order of magnitude, as well as reduce the group velocity.}
			The introduction of a zigzag has a striking effect: the bands flatten out, and more importantly, the gap size increases more than an order of magnitude.
			This increase happens because the modes where the gap is the smallest when the momentum almost entirely focuses along the strip ($k_x=k_F$), are cut off in a zigzag geometry.
			There is also a significant reduction of the group velocity where in Fig.~\ref{fig:bandstuctures} (c) the bands are nearly completely flat.
			The size of the gap, together with the velocity, directly relate to the Majorana coherence length (Eq.~\eqref{eq:majorana_coherence_length}), which we discuss in the following section.

			\comment{The introduction of an SN interface at many angles also increases performance.}
			Finally, in the presence of normal reflection, due to the availability of NS interfaces at a wide range of slopes, a zigzag system has an increased effective transmission.
			In a zigzag system, the normally reflected part of a wavefunction will eventually intersect a surface with which it is nearly perpendicular to.
			The transparency of the NS interface thus improves for non-transversal modes.

@@

			\comment{We calculate the lowest eigenfunction and plot the density.}
			Using the tight-binding package Kwant~\cite{groth_kwant:_2014}, we model a finite system to compute the Majorana wave function density for different geometries; ribbon, zigzag, parallel curve, and vertically offset sinusoids.
			In figure~\ref{fig:wavefunctions} displays the wavefunctions superimposed upon their respective geometry including the corresponding Majorana energies $E_M$, as well as $E_1$: the energies corresponding to the next lowest state.

			\begin{figure}[!htb]
			\centering
			\includegraphics[width=0.55\columnwidth]{figures/wavefunctions}
			\caption{Wavefunctions for different sizes and geometries.
			With (a) a straight system, (b) a zigzag system, (c) a system where lines parallel to a sinusoid defines the normal region, and (d) a system with vertically offset sinusoids.
			Inside the figure, we indicate the Majorana length (or coherence length) $\lambda_\textrm{M}$ and the topological energy gap $E_\textrm{gap}$.
			We observe that $\xi$ for (a) is orders of magnitude longer and $E_\textrm{gap} = E_\textrm{1} - E_M$ orders of magnitude smaller than for (b), (c), and (d), meaning that the details of the geometry do not matter for the improvements to occur.
			\label{fig:wavefunctions}}
			\end{figure}

			\comment{In a straight system, the Majoranas are very poorly localized.}
			For the ribbon system [Fig.~\ref{fig:wavefunction}(a)], we see that the decay of the density is long compared to the system size.
			We see that the wavefunction extends to the centre of the system, not showing the delocalization sought after.
			Apart from the density, as discussed in chapter \ref{chap:theory}, an overlap of Majoranas leads to a non-zero energy of the state.
			Taking this into consideration, we see that the energy of the Majorana state in the straight system is only barely below the next lowest lying eigenstate: the Majoranas are very poorly localized and topological protection against perturbations is minimal.

@@

			\comment{We calculate the topological phase diagram using the gap size.}
			Due to the size of the system, we cannot use full spectrum methods (the Pfaffian) to compute the topological invariant of our system.
			Instead, we calculate the gap by finding the absolute minimum of the spectrum $E_\textrm{gap}=\min{|E(k)|}$.
			Noting the gap closings, and that we know the topological phase for the non-zigzag system (e.g. for magnetic field vs. phase difference it is diamond shaped: Fig.~\ref{fig:pientka_phase_diagram}), we can infer the topology of the system with zigzag modulation.
			In figure ~\ref{fig:phasediagrams}, we plot the energy gap as a function of magnetic field, chemical potential, and the superconducting phase difference for both a straight system [(a) and (b)] and a zigzag system [(c) and (d)].

			\begin{figure}[!htb]
			\centering
			\includegraphics[width=0.75\columnwidth]{figures/phasediagrams}
			\caption{Phase diagrams of a straight system (a), (b) and zigzag system (c), (d), where (a), (c) are $E(\mu, B)$ and (b), (d) are $E(\phi, B)$.
			We use a generalized eigenvalue problem to find all phase boundaries at once and find the minimal energy gap by finding the minimum in the spectrum: $\min{E(k)}$.
			\label{fig:phasediagrams}}
			\end{figure}

			\comment{The phase diagram does not change much, except we see a cleaner spectrum as a result of the D class symmetry. }
			Similar to what is described by Pientka et. al~\cite{pientka2017topological} (see Fig.~\ref{fig:pientka_phase_diagram}), we see that the straight geometry, as a function of superconducting phase difference and Zeeman field (b), has a diamond-shaped phase diagram.
			The additional gap closings visible within the diamond are because of the presence of a chiral symmetry which classifies the system as BDI and causes the region of interest, where there is a single pair of Majoranas, to be relatively small.

@@

	\section{Finding the mechanism responsible for the increased gap}
		\comment{We compare different techniques to calculate the band structure of a straight system.}
		We verify the validity of the analytic model described in chapter \ref{chap:methods} and compute the bandstructure of a straight system in a tilted magnetic field.
		In figure \ref{fig:spectrum_calculation_comparison}, we compare the analytical model with band structures obtained using three different methods, implemented in Kwant.
		To show that the model has been correctly implemented, we compute the Kwant scattering matrix of the identical model used in the analytics and substitute it into the bound state condition\eqref{eq:bound_state_condition_sjl} (labelled Kwant without SOI SJL).
		To show the validity of neglecting spin-orbit in the transverse direction, we also compute the scattering matrix including the additional term and again use the bound state condition to obtain the spectrum (labelled Kwant with SOI SJL).
		We also perform the full calculation of the spectrum by computing the bandstructure of the corresponding Kwant system (labelled Full ribbon Kwant system): this result serves as the benchmark for the other techniques.

@@

			When comparing the scattering matrix approach to the full bandstructure calculation, we see that it holds only approximately: the shape of the bands is similar, but the magnitude differs significantly.
			We attribute this to the loose compliance with the short junction limit: one needs both a long superconducting coherence length, as well as the energies in the spectrum to be near zero.
			As we find the accuracy of the analytical spectrum to be relatively weak, we continue estimation of the gap in a zigzag system using the spectrum calculated with the full Kwant system.

		\subsection{Estimation of the gap through momentum cutoff}
			As explained in chapter \ref{chap:methods}, we estimate the topological gap using a quasiclassical estimate.
			We keep the ratio $4\dfrac{z_x}{z_y}$ constant and compute the gap as a function of $z_y$.
			In figure \ref{fig:quasiclassical_approximation}, the result is shown, comparing numerical gap calculations of the full zigzag system (full lines), with the result of the quasiclassical estimation.
			Note that the quasi-classics are only valid once the zigzag cuts off all trajectories which extend infinitely.
			Also important to note is that for the quasiclassical estimate, we set the longest trajectory to be the trajectory that does not include the corners, as depicted in Fig.~\ref{fig:longest_trajectory_wo_corners}

			\begin{figure}
			\centering
			\includegraphics[width=0.45\columnwidth]{images/longest_trajectory_wo_corners}
			\caption{Diagram depicting quasiclassical estimate of a zigzag system, but now with the longest trajectory chosen to avoid the corners of the zigzag.}
			\label{fig:quasiclassical_approximation}
			\end{figure}

			\begin{figure}[!htb]
			\centering
			\includegraphics[width=0.95\columnwidth]{figures/z_y_vs_E_gap}
			\caption{Comparison of the topological gap computed using a full zigzag system implemented in Kwant with the quasiclassical estimate.
			The ratio $4\dfrac{z_x}{z_y}$ is kept constant at three different values, and the gap is calculated for varying $z_y$.
			The full lines represent  the numerical gap calculations of the full zigzag system, the dashed lines represent the quasiclassical estimation.}
			\label{fig:quasiclassical_approximation}
			\end{figure}

			We would expect that the maximum gap improvement would occur near where the zigzag amplitude $z_y$ is large enough to cut off any angle in the system.
			This point is different for the three ratios but can be identified by where the gap is largest for the quasiclassical estimate (dashed lines).
			Increasing $z_y$ further would only allow longer trajectories, which would seem extra undesirable due to the magnetic field not being aligned with the diagonal sections.

@@

			The quasiclassical estimate, indicated by the dashed lines, seems to underestimate the gap consistently.
			We also see that the gap declines monotonically with respect to the zigzag amplitude, which is by design: increasing the zigzag amplitude corresponds to increasing the domain of permitted trajectories, and thus the minimum can only decrease.
			Even so, we see that the decline of the gap with respect to $z_y$ is much quicker than with an actual zigzag.
			A possible reason that the quasiclassical model sketches a darker picture than full zigzag, is because in that model, long trajectories also correspond to grazing angles of incidence at the NS interface.

>>>>

% -*- root: ./report.tex -*-
\chapter{Results \& Discussion}\label{chap:results}
		In this chapter, we show how some properties of the system change by the introduction of a zigzag geometry.
		Additionally, we attempt to isolate the mechanism behind the improved properties by estimating the gap size through the quasiclassical model introduced in Chapter \ref{chap:methods}.

@@

			\comment{We calculate the bandstructure for varying amount of zigzag.}
			We implement a tight-binding version of the model described in Section \ref{sec:physical_picture}.
			In Figure~\ref{fig:bandstuctures}, the bandstructure for saw-toothed zigzag systems Figure~\ref{fig:zigzag_methods} with varying amplitude is displayed.
			Because the zigzag has to be modelled using a supercell, we see a folded bandstructure.
			For the sake of comparison, we take the same cell size for the ribbon structure, so the bandstructure folds in the same way.

			\begin{figure}[!htb]
			\centering
			\includegraphics[width=0.55\columnwidth]{figures/bandstructures}
			\caption{Figure of the bandstuctures corresponding to the system in Figure~\ref{fig:zigzag_methods}.
			Blue lines correspond to $\phi=0$, $B=0$, red lines to $\phi=\pi$, $B = 1$.
			The three subplots are for different amplitudes of zigzag: (a) $z_y=0$, (b) $z_y=\frac{W}{4}$, and (c) $z_y=\frac{W}{2}$, where $W=200$\si{\nm} is the junction width.
			We observe that once there are no more straight trajectories inside the junction (when $z_y=\frac{W}{2}$) the spectrum becomes most insensitive to the momentum $k_x$.
			The value of the remaining parameters are $\alpha=20$ \si{\milli \eV \nm}, $g=26$, $\meff=0.02$\si{\electronmass}, $\mu=20$ \si{\milli \eV}, $\Delta=1$ \si{\milli \eV}.
			\label{fig:bandstuctures}}
			\end{figure}

			\comment{The band structures show that modulation of the geometry increases the gap by an order of magnitude, as well as reduce the group velocity.}
			The introduction of a zigzag has a striking effect: the bands flatten out, and more importantly, the gap size increases more than an order of magnitude.
			This increase happens because the modes where the gap is the smallest when the momentum almost entirely focuses along the strip ($k_x=k_F$), are cut off in a zigzag geometry.
			There is also a significant reduction of the group velocity where in Figure~\ref{fig:bandstuctures} (c) the bands are nearly completely flat.
			The size of the gap, together with the velocity, directly relate to the Majorana coherence length (Eq.~\eqref{eq:majorana_coherence_length}), which we discuss in the following section.

			\comment{The introduction of an SN interface at many angles also increases performance.}
			Finally, in the presence of normal reflection, due to the availability of NS interfaces at a wide range of slopes, a zigzag system has an increased effective transmission.
			In a zigzag system, the normally reflected part of a wavefunction will eventually intersect a surface to which it is nearly perpendicular.
			The transparency of the NS interface thus improves for non-transversal modes.

@@

			\comment{We calculate the lowest eigenfunction and plot the density.}
			Using the tight-binding package Kwant~\cite{groth_kwant_2014}, we model a finite system to compute the Majorana wave function density for different geometries namely, ribbon, zigzag, parallel curve, and vertically offset sinusoids.
			In Figure~\ref{fig:wavefunctions} displays the wavefunctions superimposed upon their respective geometry including the corresponding Majorana energies $E_M$, as well as $E_1$: the energies corresponding to the next lowest state.

			\begin{figure}[!htb]
			\centering
			\includegraphics[width=0.55\columnwidth]{figures/wavefunctions}
			\caption{Wavefunctions for different sizes and geometries.
			With (a) a straight system, (b) a zigzag system, (c) a system where lines parallel to a sinusoid defines the normal region, and (d) a system with vertically offset sinusoids.
			Inside the figure, we indicate the Majorana length (or coherence length) $\lambda_\textrm{M}$ and the topological energy gap $E_\textrm{gap}$.
			We observe that $\xi$ for (a) is orders of magnitude longer and $E_\textrm{gap} = E_\textrm{1} - E_M$ orders of magnitude smaller than for (b), (c), and (d), meaning that the details of the geometry do not matter for the improvements to occur.
			The value of the remaining parameters are $\alpha=20$ \si{\milli \eV \nm}, $g=26$, $\meff=0.02$\si{\electronmass}, $\mu=20$ \si{\milli \eV}, $\Delta=1$ \si{\milli \eV}, $B=1$\si{\tesla}, $\phi=2\pi$.
			}\label{fig:wavefunctions}
			\end{figure}

			\comment{In a straight system, the Majoranas are very poorly localized.}
			For the ribbon system Figure~\ref{fig:wavefunctions}(a), we observe that the decay of the density is long, compared to the size of the system.
			We see that the wavefunction extends to the centre of the system, not showing the delocalization sought after.
			Apart from the density, as discussed in Chapter \ref{chap:theory}, an overlap of Majoranas leads to a non-zero energy of the state.
			Taking this into consideration, it is clear that the energy of the Majorana state in the straight system is only barely below the next lowest lying eigenstate: the Majoranas are very poorly localized and the topological protection against perturbations is minimal.

@@

			\comment{We calculate the topological phase diagram using the gap size.}
			Due to the size of the system, we cannot use full spectrum methods (the Pfaffian) to compute the topological invariant of our system.
			Instead, we calculate the gap by finding the absolute minimum of the spectrum $E_\textrm{gap}=\min{|E(k)|}$.
			Noting the gap closings, and that we know the topological phase for the non-zigzag system (e.g. for magnetic field vs. phase difference it is diamond shaped: Figure~\ref{fig:pientka_phase_diagram}), we can infer the topology of the system with zigzag modulation.
			In Figure~\ref{fig:phasediagrams}, we plot the energy gap as a function of magnetic field, chemical potential, and the superconducting phase difference for both a straight system ((a) and (b)) and a zigzag system ((c) and (d)).

			\begin{figure}[!htb]
			\centering
			\includegraphics[width=0.75\columnwidth]{figures/phasediagrams}
			\caption{Phase diagrams of a straight system (a), (b) and zigzag system (c), (d), where (a), (c) are $E(\mu, B)$ and (b), (d) are $E(\phi, B)$.
			We numerically compute the dispersion of the translationally invariant versions of a straight and zigzag system and take the minimum  $\min{E(k)}$ to obtain the energy gap.
			The value of the remaining parameters are $\alpha=20$ \si{\milli \eV \nm}, $g=26$, $\meff=0.02$\si{\electronmass}, $\Delta=1$ \si{\milli \eV}.
			\label{fig:phasediagrams}}
			\end{figure}

			\comment{The phase diagram does not change much, except we see a cleaner spectrum as a result of the D class symmetry. }
			Similar to what is described by Pientka et. al~\cite{pientka_topological_2017}, we see that the straight geometry, as a function of superconducting phase difference and Zeeman field (Figure \ref{fig:phasediagrams}(b)), has a diamond-shaped phase diagram.
			The additional gap closings visible within the diamond are due to the presence of a chiral symmetry which classifies the system as BDI and causes the region of interest, where there is a single pair of Majoranas, to be relatively small.

@@

	\section{Finding the mechanism responsible for the increased gap}
		\comment{We compare different techniques to calculate the band structure of a straight system.}
		We verify the validity of the analytic model described in Chapter \ref{chap:methods} and compute the bandstructure of a straight system in a tilted magnetic field.
		In Figure~\ref{fig:spectrum_calculation_comparison}, we compare the analytical model with band structures obtained using three different methods, implemented in Kwant.
		To show that the model has been correctly implemented, we compute the Kwant scattering matrix of the identical model used in the analytics and substitute it into the bound state condition (Equation~\eqref{eq:bound_state_condition_sjl}) (labelled Kwant without SOI SJL).
		In order to prove the validity of neglecting spin-orbit in the transverse direction, we also compute the scattering matrix including the additional term and again use the bound state condition to obtain the spectrum (labelled Kwant with SOI SJL).
		We also perform the full calculation of the spectrum by computing the bandstructure of the corresponding Kwant system (labelled Full ribbon Kwant system): this result serves as the benchmark for the other techniques.

@@

			When comparing the scattering matrix approach to the full bandstructure calculation, we see that it holds only approximately: the shape of the bands is similar, but the magnitude differs significantly.
			We attribute this to the loose compliance with the short junction limit: one needs both a long superconducting coherence length, as well as the energies in the spectrum to be near zero.
			As we find the accuracy of the analytical spectrum to be relatively weak, we the continue estimation of the gap in a zigzag system using the spectrum calculated with the full Kwant system.

		\subsection{Estimation of the gap through momentum cutoff}
			As explained in chapter \ref{chap:methods}, we estimate the topological gap using a quasiclassical estimate (see Eqs.~\eqref{eq:definition_gap} and \eqref{eq:restriction_momentum}).
			We keep the ratio $\dfrac{z_x}{z_y}$ constant and compute the gap as a function of $z_y$.
			In Figure~\ref{fig:quasiclassical_approximation}, the result is shown, comparing numerical gap calculations of the full zigzag system (full lines), with the result of the quasiclassical estimation.
			Note that the quasi-classics are only valid once the zigzag cuts off all trajectories which extend infinitely.
			It is also important to notice that for the quasiclassical estimate, we set the longest trajectory to be the trajectory that does not include the corners, as depicted in Figure~\ref{fig:quasiclassical_approximation}.

			\begin{figure}
			\centering
			\includegraphics[width=0.75\columnwidth]{images/longest_trajectory_wo_corners}
			\caption{Diagram depicting quasiclassical estimate of a zigzag system, but now with the longest trajectory chosen to avoid the corners of the zigzag.}
			\label{fig:quasiclassical_approximation}
			\end{figure}

			\begin{figure}[!htb]
			\centering
			\includegraphics[width=0.95\columnwidth]{figures/z_y_vs_E_gap}
			\caption{Comparison of the topological gap computed using a full zigzag system implemented in Kwant with the quasiclassical estimate.
			The ratio $\dfrac{z_x}{z_y}$ is kept constant at three different values, and the gap is calculated for varying $z_y$.
			The full lines represent  the numerical gap calculations of the full zigzag system, the dashed lines represent the quasiclassical estimation.}
			\label{fig:quasiclassical_approximation}
			\end{figure}

			We would expect that the maximum gap improvement would occur near where the zigzag amplitude $z_y$ is large enough to cut off any angle in the system.
			This point is different for the three ratios but can be identified by where the gap is largest for the quasiclassical estimate (dashed lines).
			Increasing $z_y$ further would only allow longer trajectories, which would seem particularly undesirable due to the magnetic field not being aligned with the diagonal sections.

@@

			The quasiclassical estimate, indicated by the dashed lines, seems to underestimate the gap consistently.
			We also see that the gap declines monotonically with respect to the zigzag amplitude, which is by design: increasing the zigzag amplitude corresponds to increasing the domain of permitted trajectories, and thus the minimum can only decrease.
			Even so, we observe that the decline of the gap with respect to $z_y$ is much quicker than with an actual zigzag.
			A possible reason that the quasiclassical model sketches a darker picture than full zigzag, is because in that model, long trajectories also correspond to grazing angles of incidence at the NS interface.
